---
title: "ggplot2 intro"
author: Artemis Louyakis
email: artemis.louyakis@uconn.edu
output: html_notebook
date: 2019.01.11
---

#### First install the ggplot2 package and load it
#### You can also check your working directory and set it to a different location if needed
#### In an R notebook, your working directory is automatically the location you save this document to
#### You'll have to reset the working directory for each chunk if it is different than where you saved, or you can change the overall working directory in a "setup chunk"
```{r}
### install.packages("ggplot2")
library(ggplot2)
head(gapminder)
library(gapminder)
## find working directory
getwd()
## to change working directory (remove hash infront of code - I use a hash to block code from running if I don't need it)
#setwd("./data/")
```

#### Make plots with ggplot! 
```{r}
## basic scatter plot
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()

## add some color and make it a line graph
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             by = country, color = continent)) +
  geom_line() 

## answer to socrative question
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()

## start adding some layers
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             by = country)) +
  geom_line(aes(color = continent)) +
  geom_point(aes(color = "blue"))
```

#### Make more plots in ggplot!
```{r}
## back to scatter (no line)
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             color = continent)) +
  geom_point()

## change scale
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             color = continent)) +
  geom_point(alpha = 0.5) +
  scale_x_log10()

## add regression line
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             color = continent)) +
  geom_point(alpha = 0.5, aes(shape = continent), show.legend = FALSE) +
  scale_x_log10() +
  geom_smooth(method = "lm", size = 1.5)
```

#### Clean up your plot and save it (this is where you may find yourself in a rabbit hole for many hours)
```{r}
## clean up for publication
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, 
                             color = continent)) +
  geom_point(alpha = 0.5, size = 2, aes(shape = continent)) +
  scale_x_log10() +
  geom_smooth(method = "lm") +
  scale_y_continuous(limits = c(0,100), breaks = seq(0,100, by = 10)) +
  theme_bw() +
  ggtitle("Effects of per-capita GDP on Life Expectancy") +
  xlab("GDP per Capita ($)") +
  ylab("Life Expectancy (yrs)") +
ggsave(file = "life_expectancy.png", width = 20, height = 20, units = "cm")
ggsave(file = "life_expectancy.pdf", width = 20, height = 20, units = "cm")
```

```{r}
##socrative
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()
ggsave(file = "life_expectancy.pdf")
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()
ggsave(file = "life_expectancy.pdf")
```

#### Just for fun, let's make it a box plot
```{r}
## boxplot
plot <- ggplot(data = gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot()

##print plot
plot

##add to plot and print out 
plot + geom_point(alpha = 0.5, color = "tomato")
plot + geom_jitter(alpha = 0.5, color = "tomato")

##add to plot and make each an object
plot_tomato <- plot + geom_point(alpha = 0.5, color = "tomato")
plot_jitter <- plot + geom_jitter(alpha = 0.5, color = "tomato")

##print out your new plot objects
plot_tomato
plot_jitter
```

#### Just for fun and to up your R-game, try making your plots interactive! 
#### When in an rnotebook, you can render interative plots to html and send to all your friends!
```{r}
## To reach your A-game, make your plot interactive!
install.packages("plotly")
library(plotly)
ggplotly(plot_jitter)
## now scroll over the points on your plot and you'll see info about each one
```

